\documentclass[12pt,letterpaper]{article}

%% paper geometry
\usepackage[margin=1in]{geometry}
%% allows for including graphics
\usepackage{graphicx}
%% allows for subfigures in a figure, with their own captions
\usepackage{subfigure}
%% allows for math script fonts 
\usepackage{mathrsfs}
%% bold math
\usepackage{bm}
%% allows for line numbering during editing
\usepackage{lineno}
%% allow for page layout frames to be shown - normally commented out
%\usepackage{showframe}

% Long table support
\usepackage{longtable}

%% allows for header customization
\usepackage{fancyhdr}

% allows for colored text   %%%FIO 2 Sept 2015
\usepackage{color}

% allows for sorting cites   %%%FIO 7 Sept 2015
%\usepackage{citesort} %For some reason this package breaks compilation

% Allows for small TOCs to be made for some sections
\usepackage{shorttoc}

% Allows for changes to bullets in itemized lists, etc.
\usepackage{enumitem}

%% fonts
%%\renewcommand{\familydefault}{\sfdefault}

% Allow block commenting
\usepackage{comment}


% allows for bookmarks and hyperlinks in PDF file
\usepackage{hyperref}


%%%%%%%%%%%
%%%%%%%%%%% Document
%%%%%%%%%%% 

\begin{document}

%%% page style
%\pagestyle{headings}
\pagestyle{fancy}


\pagestyle{plain}
\label{sec:coverpaqe}



%% Title

\title{{\large SCGSR2018}}

\author{Chris Milke (SMU), Michael Kagan (SLAC),\\
  Ariel Schwartzman (SLAC), Stephen Sekula (SMU)}


%% Narrative

%VBF is important,
%but not easy
%Can be improved with HGTD, but that has its own issues
%all of this can be solved with Machine Learning
\section*{Background}
    Vector Boson Fusion (VBF) is a process of great importance to the study of Higgs boson production, with subsequent decay to the $b\overline{b}$ final state\cite{Aaboud:2018gay}, at the Large Hadron Collider (LHC). While its cross section is over an order of magnitude smaller than the leading production mechanism, gluon fusion\cite{deFlorian:2016spz}, VBF has a much more distinctive signature. As such, VBF events are far easier to identify, making them a top candidate for probing Higgs physics as well as possible physics beyond the standard model. Nonetheless, the VBF process is not without difficulty, as the signal relies on solving a difficult combinatorics problem. To date, the most significant direct observation of the Higgs and bottom quark interaction has been through the Vector Boson-Associated (VH) production of Higgs\cite{Aaboud:2018zhk}\cite{Sirunyan:2018kst}. Unfortunately, VH processes are the third largest production mechanism for Higgs Bosons, over a factor of two less than VBF. A method by which to measure VBF cleanly is therefore highly sought, as doing so would allow for independent confirmation of the VH observation, and for a more robust Higgs study overall.  
    
    While traditional identification methods struggle to tag the various jets associated with VBF efficiently and with high purity, the complications of the process lend themselves to a solution by machine learning techniques, which are adept are handling complex assortments of data. As an added measure to combat the effects of pileup (multiple proton-proton interactions in the same beam crossing) with ATLAS, VBF identification could also make use of a planned upgrade for ATLAS called the High Granularity Timing Detector (HGTD)\cite{Collaboration:2623663}. The HGTD provides timing information for tracks and vertices, allowing vertices to be distinguished in both space and time. When combined, machine learning techniques and the HGTD could make for a highly efficient method of tagging VBF processes. The overarching goal of this project then, is to develop a machine learning based-algorithm designed to enhance the identification of VBF processes within the ATLAS experiment, which can also incorporate timing information from the HGTD.  


    %This is why VBF is hard
        %two jets produced by key event, but many more present in detector. Need to identify correct jet pair
    %Use directed machine learning technique to solve these problems
        %Instead of training nn to identify vbf generically, train it to identify key aspects of vbf separately 
\section*{Vector Boson Fusion and Machine Learning}
    The VBF process is notoriously difficult to tag, as the primary process needs to be identified by relating two pairs of jets to each other. Even with no other process in the detector, the combinatorics of pairing the four jets into two pairs is already challenging. This assumes of course that only four jets are produced, which is not necessarily true; radiation from the VBF initial or final states can produce additional jets that further complicates identification. All of these issues are drastically exacerbated by the fact that the primary process is in reality never isolated. In a typical interaction, the VBF process is accompanied by a large amount of pileup and many more unrelated jets, which then make the task of identifying and pairing the correct four jets much harder. While techniques to identify VBF events exist, they are not as efficient as they could be, and cannot be easily extended to include new information about the process. Yet, such a complex pattern is a task for which machine learning techniques are ideally suited.


    The ultimate goal in this project is to construct a machine learning algorithm that can identify the signature jet pattern produced in a VBF process. In general, machine learning techniques are specially suited to handle pattern recognition, but the overwhelming complexity of the VBF process would provide a rich challenge to machine learning techniques. Taking inspiration from Louppe et al.\cite{Louppe:2017ipp}, one could imagine training a recurrent neural network with built-in knowledge of the sequential ordering of the underlying physics process. In the same sense that words in a sentence must come in a specific order to make sense, the neural network could be trained to recognize which combinations of jets make physical sense to pair together. Identifying VBF events with such a method could make a machine learning algorithm much more efficient, and provide a viable way to study Higgs production on a much larger scale than before.

    
    %As an additional measure to improve the efficiency of the search, the VBF algorithm will use a design which incorporates a high level understanding of VBF physics directly.
    
    %The idea of this design is to build the theoretical structure of the VBF process into the algorithm, so that instead of identifying the VBF as a whole, the algorithm can identify the key aspects of the VBF (e.g. specific jet pairs) separately.



    %here's how it can help with VBF 
    %but might be tricky to use.
    %Best way to handle it may be ML,
    %hence why it's perfect for use with VBF NN. 
\section*{High Granularity Timing Detector}
    The High Granularity Timing Detector (HGTD) is a subdetector planned for installation in ATLAS during the Phase-II upgrade. Its primary purpose is to provide precision (30 pico-second resolution) timing information of tracks in the forward region (2.4 < |eta| < 4.0) of ATLAS. Such timing information will be critical in the context of VBF searches within Phase-II ATLAS, where pileup can lead to vertices very near each other. Currently, these vertices can still be spatially distinguished. In Phase-II though, with up to 200 proton-proton interactions per bunch crossing, these pileup vertices will begin to partially or completely overlap each other in space. With the addition of the HGTD though, overlapping vertices will be distinguishable by their time-stamp, making them much easier to associate with the correct track. 

    The HGTD provides an advantage especially to VBF searches, as the HGTD covers the forward region, where most pileup jets and low-momentum QCD tracks end up, and also where the VBF jets go. The HGTD's presence in this region then will be a great help to disentangling the VBF jets. One notable issue with the HGTD though, is the presence of tracker services material directly in front of it. This material causes scattering of particles entering the HGTD, and can lead to difficulty in determining which hit in the HGTD should be associated with which vertices in the tracker. Once again, machine learning techniques may provide the best way to resolve this issue. Herein lies the advantage of combining research in VBF tagging with research on the HGTD. Since the VBF machine learning algorithm can benefit from the HGTD, and the HGTD can benefit from machine learning, it only seems appropriate to directly include information from the HGTD into the VBF machine learning algorithm. Doing so would make the best possible use of the HGTD's capabilities, and in the process significantly increase the efficiency of the VBF tagging algorithm.



    %SLAC is a center for machine learning research, with many experts in the field
    %slac has powerful computing resources, which would be valuable in these studies
    %HGTD and VBF synergize well as projects
    %HGTD is great for VBF study, and VBF is a great testing ground for using HGTD.
    %SLAC is ideal then b/c they have ALTIROC1, a prototype of HGTD.
    %So I can do ML work on VBF, while getting hands on understanding of what hgtd can do (with actual testing) and how it works.
    %At the same time, I can use my work on the vbf ml algorithm as feedback to steer development of hgtd.
\section*{Synergies Between SLAC and SMU}
    SLAC and SMU provide the ideal research environment for this particular project for several reasons. First among these is that both SLAC and SMU are host to some the leading practitioners of machine learning as applied to high energy physics, as well as in the field of Higgs and heavy flavor physics. This combined wealth of knowledge would provide a strong community of mentors and experts, which would be necessary for steering the development of these projects. As an additional point, it is worth noting that SLAC and SMU have already previously demonstrated a successful partnership in Higgs and bottom quark physics, as apparent in the highly Lorentz-boosted $H \to b\overline{b}$ analysis which will be presented at the Higgs Couplings Conference in November 2018. This analysis was initiated in part by a collaboration between SLAC and SMU, and can serve as a foundation upon which to build future success in this project.

    In addition to its experts in machine learning and Higgs physics, SLAC also has a range of experts on distributed systems and massively parallel programming techniques. Such systems are necessary for training machine learning algorithms over large data sets, as would be the case for the algorithm proposed here. But not only does SLAC have experts on distributed systems, but SLAC is home to a vast, growing array of massively parallel processors and GPUs. These GPU clusters would serve as a crucial resource during the training of the machine learning algorithm proposed here.
    
    The most important reason for collaborating with SLAC however, is that SLAC is currently developing a prototype of the HGTD, known as ALTIROC1. Having access to a physical prototype of the HGTD would be invaluable to the development of a machine learning algorithm incorporating HGTD information. Instead of having to merely use simulated data, the presence of ALTIROC1 at SLAC would allow for genuine testing data and an assessment of what information from the HGTD would really look like. Furthermore, development of the VBF algorithm could be used to inform the ongoing development of the HGTD. Undoubtedly, situations will arise in which the information ALTIROC1 provides is found to be less than ideal for actual use in identifying physics processes. By using ALTIROC1 data for the development of a VBF algorithm at SLAC, these deficiencies can be identified during development, and passed directly to the HGTD design team. 




    %research done here is immediately helpful to vbf studies, but extends beyond them. 
    %the ml techniques used here for two jets for vbf can be used for other physics events, and extended to more complex processes of more jets in the future.
    %This can aid in the development of the hardware of the hgtd,
    %and can act as a proof of concept of how hgtd output can be used via ml algorithms
\section*{Looking Forward}
    The immediate goal of this research project is to develop a machine learning algorithm, utilizing the HGTD, to identify vector boson fusion events within ATLAS in service of a more detailed study of the Higgs and bottom quark coupling. Yet, the effects of this research will extend far beyond VBF studies alone. The combination of modern and advanced machine learning techniques with this key physics process provides an opportunity for new advances and discoveries. The work done here with VBF processes could thus be extended to a wide range of additional event types in later work. Additionally, the use of prototype HGTD information in this algorithm, and the assessment of that information, has the potential to influence the overall development of the HGTD leading to its installation in Phase-II. This algorithm can act as a testing ground for key features of the HGTD, and issues identified here will benefit any project in the future which utilizes HGTD hardware. As a final point, not only will this algorithm's development help to steer the HGTD's hardware design, but it also serves as a template for how to use the HGTD's output in software. The most efficient way to use HGTD timing information may very well be through machine learning, and the techniques incorporating the HGTD in this project can be used as an example for any other machine learning project in the future. Overall, while the goal of this project is fairly narrow, the effects of its development are broad and far reaching.




%% Appendix, if needed
\newpage
\begin{appendix}
\end{appendix}


%% Bibliography

\bibliographystyle{JHEP}
\bibliography{project}



\end{document}



